---
title: "Algorithms"
output: html_document
date: '2022-05-10'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Simulation Setup
```{r}
library(sampling)
library(survey)
library(dplyr)

#number of sites
n <- 100
#number of possible sites
N <- 1000

set.seed(12412)

simdf <- data.frame(id = 1:N, x = runif(N), y = runif(N))
yobs <- (exp((simdf$x-0.5)^2)+exp((simdf$y-0.5)^2))
yobs <- 100-((yobs-min(yobs))/(max(yobs-min(yobs))))*100+
+ (rnorm(N)+5)*5
ypps <- exp(yobs/10) #continuous variable
q1obs <- sample(1:3,N,replace = TRUE) #categorical variable
q2obs <- sample(LETTERS[1:2], N, replace = TRUE) #categorical variable
simdf <- cbind(simdf,yobs,ypps,q1obs,q2obs)

#distance matrix for candidate sites
dist_mat <- as.matrix(dist(simdf %>% select(x, y)))
```

# GRST Sampling
Using the vingette example https://cran.r-project.org/web/packages/spsurvey/vignettes/sampling.html
```{r}
library(spsurvey)

simdf_sf <- st_as_sf(simdf, coords = c("x", "y"), crs = 4326)
simdf_sf <- st_transform(simdf_sf, crs = 5070)

#grts sample where all sites have equal probability of selection
eqprob <- grts(simdf_sf, n_base = n)
spbalance(dist_mat, rep(n/N,N), eqprob$sites_base$id)

#plot against possible sites
sp_plot(eqprob, simdf_sf, key.width = lcm(3))
```

Unequal probabilites based on a categorical variable (diff sample number between categories).
```{r}
#named vector of sample sizes for categories
caty_n <- c("1" = 30, "2" = 10, "3" = 10)

uneqprob <- grts(
  simdf_sf %>% mutate(q1obs = as.character(q1obs)),
  n_base = 50,
  caty_var = "q1obs",
  caty_n = caty_n
)

sp_plot(uneqprob, simdf_sf, key.width = lcm(3))
```

Inclusion probability proportional to a continuous variable
```{r}
propprob <- grts(
  simdf_sf,
  n_base = 50,
  aux_var = "ypps"
)

sp_plot(propprob, simdf_sf, key.width = lcm(3))
```
## Stratification

Equal inclusion probabilities within stratum
```{r}
strata_n <- c(A = 25, B = 15)
strat_eqprob <- grts(simdf_sf, n_base = strata_n, stratum_var = "q2obs")

#plots broken down by strata
sp_plot(
  strat_eqprob,
  formula = siteuse ~ q2obs,
  simdf_sf,
  key.width = lcm(3)
)
```
Unqeual inclusion probabilities within stratum
```{r}
caty_n <- list(
  A = c("1" = 15, "2" = 5, "3" = 5),
  B = c("1" = 7, "2" = 5, "3" = 3)
)

strat_uneqprob <- grts(
  simdf_sf,
  n_base = strata_n,
  stratum_var = "q2obs",
  caty_var = "q1obs",
  caty_n = caty_n
)
```

Probabilities within strata proportional to continuous variable
```{r}
strat_propprob <- grts(
  simdf_sf,
  n_base = strata_n,
  stratum_var = "q2obs", #stratification categorical variable
  aux_var = "ypps" #continuous variable 
)
```

Let's try with an area sample
```{r}
set.seed(4447864)

shp_file <- st_read(here::here("data/greatvalley_outline.shp"))
eqprob_grts_area <- grts(shp_file, n_base = n)
```


# Cube Sampling
Sample sizes is random and not fixed
- include a first-order inclusion probability vector when balancing the auxiliary variables in order to enforce a sample size

Using `spsurvey::samplecube()` here, but there's also `BalancedSampling::cube()`
```{r}
#number of sites
n <- 100
#number of possible sites
N <- 1000

#plot potential sites
par(mar = c(1,1,1,1), xaxs = "i",yaxs = "i")
plot(simdf$x,simdf$y, axes = F,cex = 0.5,pch = 19, xlim = c(0,1), ylim = c(0,1))
box()

#
set.seed(200694)
pik <- rep(n/N,N) #vector of inclusion probabilities
x <- as.matrix(cbind(simdf$x,simdf$y))
bal <- samplecube(x,pik,comment=TRUE,method=1)

spbalance(dist_mat,rep(n/N,N),(1:1000)[bal==1])

## BalancedSampling version
bal_bs <- BalancedSampling::cube(pik,x)
```

# Spatially correlated Poisson sampling (SCPS)
```{r}
library(BalancedSampling)

n <- 100
N <- 1000
set.seed(200694)

p <- rep(n/N,N)
x <- cbind(simdf$x,simdf$y)

sam_scps <- scps(p,X)
length(sam_scps)

spbalance(dist_mat,rep(n/N,N),sam_scps)

par(mar = c(1,1,1,1), xaxs = "i",yaxs = "i")
plot(simdf$x,simdf$y, axes=F,cex=0.5,pch=19, xlim=c(0,1),ylim=c(0,1))
points(simdf$x[sam_scps],simdf$y[sam_scps],pch=1, cex=2)
box()
```
# Local Pivotal Method
There are two versions of the algorithm, lpm1 is better balanced, lpm2 is faster
```{r}
n <- 100
N <- 1000
set.seed(200694)

p <- rep(n/N,N)
x <- cbind(simdf$x,simdf$y)

sam_lpm1 <- lpm1(p,X)
spbalance(dist_mat,rep(n/N,N),sam_lpm1)

sam_lpm2 <- lpm2(p,X)
spbalance(dist_mat,rep(n/N,N),sam_lpm2)
```

# Algorithm Characteristics

GRTS `spsurvey::grst`:
- outputs `spdesign` class

Cube `spsurvey::samplecube`:
- outputs binary inclusion vector of length N

SCPS `BalancedSampling::scps`:
- outputs vector of length n of included site indexes

LPM `BalancedSampling::lpm1`
- - outputs vector of length n of included site indexes
