---
title: "IUCN evaluation"
output: html_document
date: '2022-05-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(sf)
library(survey)
library(spsurvey)
library(BalancedSampling)
library(ggplot2)
library(sampling)
library(raster)

devtools::load_all()
```


```{r}
#sf::sf_use_s2(FALSE)

#read in IUCN data
ranges <- st_read(here::here("data/iucn_terrestrial_mammals/MAMMALS_TERRESTRIAL_ONLY.shp")) %>%
  rename(order = "order_")

#let's get bats
bats <- ranges %>% filter(order == "CHIROPTERA") %>%
  st_make_valid()

#let's get Canada
cdn <- getData("GADM",country="can",level=0) %>%
  st_as_sf() %>%
  st_make_valid()

#crop bats to canada
cdn_bats <- st_intersection(bats, cdn)
```
Let's try just quebec
```{r}
cdn_regions <- getData("GADM",country="can",level=1) %>%
  st_as_sf()

qc <- cdn_regions %>% filter(NAME_1 == "Québec") %>%
  st_make_valid()

qc_bats <- st_intersection(bats, qc)

plot(st_geometry(cdn_bats))
```

Let's do an equal probability monitoring network for all of quebec
```{r}
qc_proj <- st_transform(qc, crs = 5070)
qc_bats_proj <- st_transform(qc_bats, crs = 5070)

set.seed(12356)
n <- 100

qc_cand_points <- get_cand_points(qc_proj, n)
qc_eqprob <- get_eqprob(qc_cand_points, n = n)
```

Evaluate cube method for species coverage
```{r}
#extract species list for each point from range maps
cube_points <- qc_cand_points %>%
  filter(id %in% qc_eqprob$cube)

#get a dense matrix of intersections between points and ranges
#points and names are brute force reassigned to the matrix, need to double check that that works
bat_points <- st_intersects(cube_points, qc_bats_proj, prepared = TRUE, sparse = FALSE) %>%
  as_tibble() %>%
  magrittr::set_colnames(qc_bats_proj$binomial) %>%
  bind_cols(cube_points)
```

Plot a range with it's supposedly intersecting points (check that intersection did what we thought)
```{r}
bat_name <- "Lasionycteris noctivagans"

bat_range <- qc_bats_proj %>% 
  filter(binomial == bat_name)

bat_name_enc <- enquo(bat_name)
bat_range_points <- bat_points %>%
  dplyr::select(id, x, presence = bat_name) %>%
  filter(presence == TRUE)

ggplot() +
  geom_sf(data = bat_range) +
  geom_sf(data = bat_range_points)
```

Now we want multiple equa probability replicates at multiple sample sizes.
```{r}
#get more candidate points, increase N by pt_density argument
qc_cand_points <- get_cand_points(qc_proj, n, pt_density = 100)

#get multiple replicates for a single sample size
repn <- 3

get_reps <- function(n, nreps) {
  purrr::map_dfr(1:nreps, 
                 function(rep) {
                   print(c(rep, n))
                   get_eqprob(qc_cand_points, n = n) %>% 
                     mutate(rep = rep)  
                 })
                 
}

sample_sizes <- c(10, 20, 50, 100, 500, 1000, 5000)
samplesize_sim <- purrr::map_dfr(sample_sizes[1:3], ~get_reps(.x, nreps = repn) %>% mutate(sample_size = .x))

```

Get species counts for each candidate point
```{r}
specieslist <- st_intersects(qc_cand_points, qc_bats_proj, prepared = TRUE, sparse = FALSE) %>%
  as_tibble() %>%
  magrittr::set_colnames(qc_bats_proj$binomial) %>%
  bind_cols(qc_cand_points)
```
Look up species lists for each monitoring network
```{r}
specieslist_long <- specieslist %>%
  pivot_longer(cols = c(-x, -id), names_to = "species") %>%
  filter(value == TRUE) %>%
  dplyr::select(-value)


design_lists <- samplesize_sim %>%
  pivot_longer(cols = c(-id, -x, -rep, -sample_size), names_to = "algorithm") %>%
  filter(value == 1) %>%
  dplyr::select(-value) %>%
  left_join(specieslist_long)

design_lists %>% 
  dplyr::group_by(rep, algorithm, sample_size) %>%
  summarize(richness = n_distinct(species))
```

