---
title: "IUCN evaluation"
output: html_document
date: '2022-05-27'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(sf)
library(survey)
library(spsurvey)
library(BalancedSampling)
library(ggplot2)
library(sampling)
library(raster)

devtools::load_all()
```


```{r}
sf::sf_use_s2(FALSE)

#read in IUCN data
ranges <- st_read(here::here("data_store/iucn_terrestrial_mammals/MAMMALS_TERRESTRIAL_ONLY.shp")) %>%
  rename(order = "order_") %>%
  st_make_valid()

#let's get bats
bats <- ranges %>% filter(order == "CHIROPTERA") %>%
  st_make_valid()

#let's get Canada
cdn <- getData("GADM",country="can",level=0) %>%
  st_as_sf() %>%
  st_make_valid()

#crop bats to canada
cdn_bats <- st_intersection(bats, cdn)
```

Get WWF world ecoregions
```{r}
ecoreg <- st_read(here::here("data_store/wwf_ecoregions/wwf_terr_ecos.shp")) 
```

Let's try just quebec
```{r}
cdn_regions <- getData("GADM",country="can",level=1) %>%
  st_as_sf()

qc <- cdn_regions %>% filter(NAME_1 == "Québec") %>%
  st_make_valid()

qc_bats <- st_intersection(bats, qc)

#get canadian mammals, use buffer to fix self-intersection errors
ranges_buff <- st_buffer(ranges, 0)
qc_mammals <- st_intersection(ranges_buff, qc)

plot(st_geometry(cdn_bats))
```

Let's do an equal probability monitoring network for all of quebec
```{r}
qc_proj <- st_transform(qc, crs = 5070)
qc_bats_proj <- st_transform(qc_bats, crs = 5070)
qc_mammals_proj <- st_transform(qc_mammals, crs = 5070)

# set.seed(12356)
# n <- 100
# 
# qc_cand_points <- get_cand_points(qc_proj, n)
# qc_eqprob <- get_eqprob(qc_cand_points, n = n)
```

Evaluate cube method for species coverage
```{r}
# #extract species list for each point from range maps
# cube_points <- qc_cand_points %>%
#   filter(id %in% qc_eqprob$cube)
# 
# #get a dense matrix of intersections between points and ranges
# #points and names are brute force reassigned to the matrix, need to double check that that works
# bat_points <- st_intersects(cube_points, qc_bats_proj, prepared = TRUE, sparse = FALSE) %>%
#   as_tibble() %>%
#   magrittr::set_colnames(qc_bats_proj$binomial) %>%
#   bind_cols(cube_points)
```

Plot a range with it's supposedly intersecting points (check that intersection did what we thought)
```{r}
# bat_name <- "Lasionycteris noctivagans"
# 
# bat_range <- qc_bats_proj %>% 
#   filter(binomial == bat_name)
# 
# bat_name_enc <- enquo(bat_name)
# bat_range_points <- bat_points %>%
#   dplyr::select(id, x, presence = bat_name) %>%
#   filter(presence == TRUE)
# 
# ggplot() +
#   geom_sf(data = bat_range) +
#   geom_sf(data = bat_range_points)
```

Now we want multiple equa probability replicates at multiple sample sizes.
```{r}
n <- 100
#get more candidate points, increase N by pt_density argument
qc_cand_points <- get_cand_points(qc_proj, n, pt_density = 100)

#get multiple replicates for a single sample size
repn <- 100

get_reps <- function(n, nreps) {
  purrr::map_dfr(1:nreps, 
                 function(rep) {
                   print(c(rep, n))
                   get_eqprob(qc_cand_points, n = n) %>% 
                     mutate(rep = rep)  
                 })
                 
}

sample_sizes <- c(10, 20, 50, 100, 500, 1000, 5000)
samplesize_sim <- purrr::map_dfr(sample_sizes[1:3], ~get_reps(.x, nreps = repn) %>% mutate(sample_size = .x))

library(furrr)
#options(future.debug = FALSE)
#have to source internal functions so the workers can see them
source("R/get_eqprob.R")
plan("multisession", workers = length(sample_sizes))
samplesize_sim <- furrr::future_map_dfr(sample_sizes, 
                                        ~get_reps(.x, nreps = repn) %>% mutate(sample_size = .x),
                                        .options = furrr_options(packages = "monitoringSim",  
                                        seed = TRUE))

```

Get species counts for each candidate point
```{r}
bat_specieslist <- st_intersects(qc_cand_points, qc_bats_proj, prepared = TRUE, sparse = FALSE) %>%
  as_tibble() %>%
  magrittr::set_colnames(qc_bats_proj$binomial) %>%
  bind_cols(qc_cand_points)

mammal_specieslist <- st_intersects(qc_cand_points, qc_mammals_proj, prepared = TRUE, sparse = FALSE) %>%
  as_tibble() %>%
  magrittr::set_colnames(qc_mammals_proj$binomial) %>%
  bind_cols(qc_cand_points)
```

Look up species lists for each monitoring network
```{r}
specieslist_long <- mammal_specieslist %>%
  pivot_longer(cols = c(-x, -id), names_to = "species") %>%
  filter(value == TRUE) %>%
  dplyr::select(-value)


design_lists <- samplesize_sim %>%
  dplyr::select(-x) %>%
  pivot_longer(cols = c(-id, -rep, -sample_size), names_to = "algorithm") %>%
  filter(value == 1) %>%
  dplyr::select(-value) %>%
  left_join(specieslist_long, by = "id")

richness_df <- design_lists %>% 
  dplyr::group_by(rep, algorithm, sample_size) %>%
  summarize(richness = n_distinct(species, na.rm = TRUE))

richness_stats <- richness_df %>%
  group_by(algorithm, sample_size) %>%
  summarize(mean = mean(richness),
            sd = sd(richness))

ggplot(richness_stats, aes(x = sample_size, y = mean, color = algorithm)) +
  geom_line() + 
  labs(y ="Mean Richness", x = "Sample Size") +
  theme_classic()
 # geom_ribbon(aes(ymin = mean - sd, ymax = mean + sd))

ggsave("figures/ex_richness_sat.jpg", width = 5, height = 3.5, units = "in")
```

Get spatial balance metric for each rep
```{r}
# design_index_df <- design_lists %>% 
#   dplyr::select(-species) %>%
#   distinct()
#   

```

Land cover evaluations
```{r}
can_lc <- raster("data_store/CAN_NALCMS_2010_v2_land_cover_30m/CAN_NALCMS_2010_v2_land_cover_30m.tif") 
  

qc_lc <- can_lc %>% #crop(can_lc, st_bbox(cdn)) %>%
  #projectRaster(., crs = 5070) %>%
  mask(., qc_proj)

plot(qc_lc)
```

