---
title: "Quebec Sim"
output: html_document
date: '2022-07-06'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(furrr)
library(sf)
library(survey)
library(spsurvey)
library(BalancedSampling)
library(sampling)

#devtools::load_all()
```

Load Data
```{r}
load(here::here("data/qc.rda"))
qc_proj <- st_transform(qc, crs = 5070)
```

Set up simulation for Quebec
```{r}
#number of points
n <- 100
#number of replicates for a single sample size
repn <- 100

#list of sample sizes
# sample_sizes <- c(10, 20, 50, 100, 200, 300, 500)

# get sample sizes on a log scale, this is a range from about 10 to 550
log_samp <- seq(2.3, 6.3, by = 0.4)
sample_sizes <- round(exp(log_samp))

# to get more candidate points (increase N), larger pt_density 
# qc_cand_points <- get_cand_points(qc_proj, n, pt_density = 100)

usethis::use_data(qc_cand_points)
```

Plot points on shapefile
```{r}
# ggplot() + geom_sf(data = qc_proj[1], fill = "grey") + 
#   geom_sf(data = qc_cand_points, fill = "black", size = 0.05)
```

Get ecoregions for stratification
```{r}
#get ecoregion data
ecoreg <- st_read(here::here("data/wwf_ecoregions/wwf_terr_ecos.shp")) %>%
  st_transform(ecoreg, crs = 5070) %>%
  st_make_valid()

# make sure CRS matches candidate points
raster::compareCRS(ecoreg, qc_cand_points)

# get the ecoregion for each point
ecoreg_pts <- st_intersection(qc_cand_points, ecoreg)

usethis::use_data(ecoreg_pts)
```

## Equal probability sampling
```{r}
# test run
# N <- dim(qc_cand_points)[1]
# samplesize_sim <- purrr::map_dfr(sample_sizes[1:2], ~get_eqprop_reps(.x, N = N, pt_df = qc_cand_points, nreps = 2) %>% mutate(sample_size = .x))

# Parallelize Version
#options(future.debug = FALSE)
#have to source internal functions so the workers can see them
source(here::here("R/get_eqprob.R"))
N <- dim(qc_cand_points)[1]

plan("multisession", workers = length(sample_sizes))
samplesize_sim <- furrr::future_map_dfr(sample_sizes, 
                                        ~get_eqprop_reps(.x, N = N, pt_df = qc_cand_points, nreps = repn) %>% 
                                          mutate(sample_size = .x),
                                        .options = furrr_options(seed = TRUE
                                        #packages = "monitoringSim"  
                                        ))  %>% 
    unite("simid", rep, sample_size, algorithm, remove = FALSE)

usethis::use_data(samplesize_sim)
```

## Stratified Equal Probability sampling

How many points are in each ecoregion? Exclude ecoregions with very few points
```{r}
inc_ecoreg_pts <- ecoreg_pts %>% 
  count(ECO_NUM) %>%
  filter(n > 10)
```

What area does each ecoregion cover?
```{r}
#clip the shapefile to quebec (study frame) to get are
ecoreg_qc <- st_intersection(ecoreg, qc_proj)
ecoreg_qc$AREA <- st_area(ecoreg_qc)

#https://stackoverflow.com/questions/69539592/calculate-the-surface-area-of-a-variable-in-a-shapefile-in-r#:~:text=Load%20your%20multipolygon%20in%20R,(row)%20in%20your%20multipolygon.&text=%3E%20df2,%3D%20%22id%22%2C%20all.

ecoreg_areas <- ecoreg_qc %>% 
  filter(ECO_NUM %in% ecoreg_pts$ECO_NUM) %>% #exclude regions with too few points
  group_by(ECO_NUM) %>% 
  summarize(st_union(geometry), 
            area_NAME = sum(AREA)) %>%
  mutate(PERC_AREA = area_NAME/sum(area_NAME))

```

Format ecoregion and point data
```{r}
ecoregion_points_data <- ecoreg_pts %>% 
  filter(ECO_NUM %in% inc_ecoreg_pts$ECO_NUM) %>% #exclude ecoregions with too few points
  arrange(id) %>%
  left_join(ecoreg_areas %>%
              st_drop_geometry() %>%
              select(ECO_NUM, PERC_AREA))

usethis::use_data(ecoregion_points_data)
```

```{r}
# test <- purrr::map_dfr(sample_sizes[1:2], ~get_strata_reps(.x, strat_df = inc_pts,
#                                                    nreps = 2, strata_col = ECO_NUM, area_col = PERC_AREA) %>% mutate(sample_size = .x))


source(here::here("R/get_eqprob_strat.R"))

plan("multisession", workers = length(sample_sizes))
strata_sim <- furrr::future_map(sample_sizes, 
                                        ~get_strata_reps(.x, strat_df = ecoregion_points_data, 
                                                    nreps = repn, strata_col = ECO_NUM, area_col = PERC_AREA), 
                                        .options = furrr_options(seed = TRUE))

ecoregion_inc_prob <- purrr::map_dfr(strata_sim, 2)
strata_sim_points <- purrr::map_dfr(strata_sim, 1) %>%
  unite("simid", rep, sample_size, algorithm, remove = FALSE)

usethis::use_data(strata_sim_points)
usethis::use_data(ecoregion_inc_prob)
```
## Simulate latitude bands

```{r}
# number of latitude bands

```


Function for plotting a single rep against the shapefile
```{r}

plot_design <- function(data, id){
  
  data %>%
    filter(simid == id) %>%
    st_as_sf()
  
  plot(qc_proj)
  points(sf::st_coordinates(qc_cand_proj)[,1],
         sf::st_coordinates(qc_cand_proj)[,2])
    
}
```

